---
title: "AS404- Assignment 01"
author: "S/16/342  D.G.M.D.Dimbulkumbura"
date: "3/24/2022"
output: pdf_document
---
---
title: "AS404 Assignment 01"
author: "S/16/342  D.G.M.D.Dimbulkumbura"
date: "3/24/2022"
output: pdf_document
---
```{r}
#Call to the tidyverse package
library(tidyverse)
library(magrittr)
library(dplyr)
```


#..........Q(a)..........
```{r}


#Read the dataset
crimeData=read_csv(file="E:/400 Level Sem 1/AS 404/Assignment/AS404_S16_342/crimeData.csv")



  
#Defining the column types

crimeData=read_csv(file="E:/400 Level Sem 1/AS 404/Assignment/AS404_S16_342/crimeData.csv",
                   col_types = c("dildiiiliiilii"))


  #rename the column names in ‘snake case’ format

crimeData=crimeData%>%
rename(crime_rate=CrimeRate,
       youth=Youth,
       southern=Southern,
       education=Education,
       expenditure_year_0=ExpenditureYear0,
       labour_force=LabourForce,
       males=Males,
       more_males=MoreMales,
       state_size=StateSize,
       youth_unemployment=YouthUnemployment,
       mature_unemployment=MatureUnemployment,
       high_youth_unemploy=HighYouthUnemploy,
       wage=Wage,
       below_wage=BelowWage)

```
#..........Q(b)..........

```{r}
crimeData%>%
  ggplot(mapping = aes(x=crime_rate))+
  geom_histogram(bins=15,color="blue",fill="yellow")+
labs(title="Distribution of Crime Rate",x="Crime Rate",y="Count")
  

```

..........Q(c)..........

```{r}
crimeData%>%
  group_by(southern)%>%
  summarise(avg_crime_rate=mean(crime_rate))


```

Southern states have high average crime rate than the non- southern states


..........Q(d)..........

```{r}
crimeData%>%
  ggplot()+
  geom_boxplot(aes(x=southern,y=crime_rate,fill=southern))+
labs(title="Crime rates in southern and non-southern states",x="Southern / Non-Southern",y="Crime Rate")
  
  
  
```
The first quartile of Non-southerns' lower than the Southerns. But median & 3rd quartile of Non-southerns' higher than the Southerns. Also Non-southerns have higher range compared to the Southerns.

..........Q(e)..........

```{r}
ggplot(crimeData, aes(x=more_males, y=crime_rate, fill=high_youth_unemploy)) + 
    geom_boxplot() +
    facet_wrap(~high_youth_unemploy)+
  ggtitle("Crime rate comparing (more_males VS high_youth_unemploy)")
  
```

..........Q(f)..........

```{r}
ggplot(crimeData, aes(x=more_males, y=youth_unemployment))+geom_bar(stat="identity",fill="blue")+
labs(title="youth unemployment of the states based on gender composition",x="Number of Males > Number of Females",y="Youth unemployment")
```

..........Q(g)..........

```{r}
ggplot(crimeData, aes(x=wage, y=crime_rate)) + 
    geom_point(color="red")+
labs(title="Relationship between the Crime rate and Median weekly wage",x="Median weekly wage",y="Crime Rate")
```
  
..........Q(h)..........  
       
```{r}
linear_mod=lm(crimeData$crime_rate~crimeData$wage)
```
    

..........Q(i)..........
    
```{r}
summary(linear_mod)
```
    
Y=(0.12722)*X+(35.97031) ;Where Y=Crime Rate & X=Median weekly wage  
